---
title: "Using `AssignSampleIDs` with Example Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using AssignSampleIDs with Example Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AssignSampleIDs)
```

## Introduction
This vignette demonstrates how to use the AssignSampleIDs package with the included test_data dataset. The test_data dataset contains example lab IDs, personal IDs, and sample collection dates.

## Loading the Example Data
The test_data dataset is included in the package and can be loaded as follows:

```{r test_data}
data(test_data)
head(test_data)
```

## Assigning Sample IDs
The assign_sample_id function assigns unique sample IDs, annotates duplicates, and tracks tumor occurrences for each patient. Here’s how to use it with the test_data dataset:

```{r assign_sample_ids}
result <- assign_sample_id(
  this_data = test_data,
  start_id = 100,
  lab_id_col = "lab_id",
  personal_id_col = "personal_id",
  date_col = "date_of_sample",
  verbose = TRUE
)

head(result)
```

## Explanation of the Output
The output of the assign_sample_id function includes the following columns:

- lab_id: The original lab ID column.
- personal_id: The original personal ID column.
- date_of_sample: The original date of sample collection column.
- sample_id: The formatted sample ID column, which includes unique IDs, replicate numbers, and tumor annotations.

## Summary Statistics
If the verbose parameter is set to TRUE, the function prints summary statistics about the data, including:

- The number of unique samples.
- The number of patients with replicates.
- The number of samples with multiple tumors.

## Conclusion
This vignette demonstrated how to use the AssignSampleIDs package with the test_data dataset. For more details, refer to the function documentation or explore other features of the package.
